---
title: "Dissertation"
output: word_document
---
## look at correlation co-efficients and regression --> maybe email kirils and check that it would be suitable? 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A longitudinal study, using wave 8 from the understanding society covid-19 data and wave ... from the main stage Understanding Society Data. 
Aggregate data of individual repsonses. 
Chosen wave 8 (collected 24/03 to 31/03 2021) as this was the end of the third national lockdown (29/03), thus would have seen people facing furlough and other economic difficulties with industries shut. 

Wave 10 of main stage data used

```{r}
# Downlaod data 
library(vroom)
library(tidyverse)
March21 <- vroom("UKDA-8644-tab/tab/ch_indresp_w.tab", 
                 col_select = c(pidp, racel_dv, ch_hsownd_cv, ch_sex_cv, ch_age, ch_newfurlough, ch_hhincome_amount, ch_ucreditb65, ch_hhnum, ch_xphs_cv, ch_ff_sempderived, ch_hours, ch_betaindin_xw, ch_betaindin_lw))

Mainstage <-  vroom("UKDA-6614-tab/tab/ukhls_w10/j_indresp.tab", 
                    col_select = c(pidp, j_racel_dv,  j_sex_dv,  j_jbisco88_cc, j_hidp, j_fimnnet_dv, j_jbsemp, j_jbhrs, j_indinub_xw)) 

Mainstage2 <- vroom("UKDA-6614-tab/tab/ukhls_w10/j_hhresp.tab", 
                    col_select = c(j_hidp, j_hsownd)) 
```



Join the waves. Need to join via pidp, then filter out people who have not had a constant occupation.
ch = covid (March 2021)
j = mainstage (2018-2019) ----- three years difference (three years between ages)

```{r}
## first join the two mainstage datasets 

Mainstage1 <- Mainstage %>%
  left_join(Mainstage2, by = "j_hidp")


## join all three --> left join will filter out any observations in the mainstage data that werent in the COVID-19 data. 

EconData <- March21 %>%
  left_join(Mainstage1, by = "pidp")

```



Recode ethnicities --> nominal variable 
```{r}
EconData <- EconData %>% 
  mutate(WhiteEthn = recode(racel_dv ,
        '1' = "White",
        '2' = NA_character_,
        '3' = NA_character_,
        '4' = NA_character_,
        '9' = "Ethnic Minority",
        '10' = "Ethnic Minority",
        '11' = "Ethnic Minority", 
        '12' = "Ethnic Minority",
        '-9' = NA_character_,
        .default = NA_character_))
```

Recode employment type --> nominal variable 

```{r}


EconData <- EconData %>% mutate(EmployedPre = recode(j_jbsemp, 
                           '1' = "Employee", 
                           '2' = "Self Employed", 
                           '-8' = NA_character_,
                           '-2' = NA_character_, 
                           '-1' = NA_character_))                         
                                                     
```

Recode universal credit 

```{r}
EconData <- EconData %>%
  mutate(UniCredit = recode(ch_ucreditb65, 
                                 '1' = "No", 
                                 '2' = "Yes", 
                                 '3' = "Yes", 
                                 '4' = "Yes", 
                                 '5' = "Yes", 
                                 '6' = "Yes",
                            '-9' = NA_character_, 
                                 '-8' = NA_character_, 
                                 '-2' = NA_character_, 
                                 '-1' = NA_character_))

EconData <- EconData %>%
  filter(!is.na(UniCredit))

EconData <- EconData %>%
      mutate(UniBin = ifelse(UniCredit == "Yes", 0, 1))
```




Recode working hours to show difference in working hours pre and during covid. 

```{r}
EconData <- EconData %>%
  mutate(HourDif = ch_hours - j_jbhrs)

EconData <- EconData %>%
  filter(HourDif < 100)
```


Recode housing payments and whether they are up to date 

```{r}
EconData <- EconData %>%
  filter(ch_xphs_cv %in% 1:2) %>%
      mutate(Payments = ifelse(ch_xphs_cv == 1, "Yes", "No"))

EconData <- EconData %>%
      mutate(PaymentsBin = ifelse(Payments == "Yes", 0, 1))
```

Create  new variable that contains: 
 # 1 = EM and employee
 # 2 = Em and self-employed
 # 3 = White and employee 
 # 4 = White and self-employed 
 
```{r}
EconData <- EconData %>%
  filter(!is.na(WhiteEthn)) %>%
  filter(!is.na(EmployedPre)) %>%
  mutate(IndepVar  = case_when(
    WhiteEthn == "Ethnic Minority" & EmployedPre == "Employee" ~ 1, 
    WhiteEthn == "Ethnic Minority" & EmployedPre == "Self Employed" ~ 2, 
    WhiteEthn == "White" & EmployedPre == "Employee" ~ 3, 
    WhiteEthn == "White" & EmployedPre == "Self Employed" ~ 4
  ))

EconData <- EconData %>%
  filter(!is.na(WhiteEthn)) %>%
  filter(!is.na(EmployedPre)) %>%
  mutate(IndepVar2  = case_when(
    WhiteEthn == "Ethnic Minority" & EmployedPre == "Employee" ~ "Ethnic Minority Employee", 
    WhiteEthn == "White" & EmployedPre == "Employee" ~ "White Employee", WhiteEthn == "Ethnic Minority" & EmployedPre == "Self Employed" ~ "Ethnic Minority Self-Employed",
    WhiteEthn == "White" & EmployedPre == "Self Employed" ~ "White Self-Employed"
  ))
EconData <- EconData %>% filter(!is.na(ch_betaindin_lw))
```







###########################################################################################################










Descriptive statistics 

```{r}


# Categrorical varaibles 
EconData %>%
  group_by(UniCredit) %>%
  count() 

EconData %>%
  group_by(IndepVar) %>%
  count() 

EconData %>%
  group_by("Payments") %>%
  count() 

# Continuous variables 

  mean(EconData$HourDif)

  median(EconData$HourDif)

  min(EconData$HourDif)

  max(EconData$HourDif)
  
  sd(EconData$HourDif)
EconData <- EconData %>% mutate(IndepVar = as.factor(IndepVar))


#######

EconData %>%
  count(WhiteEthn)

EconData %>%
  count(EmployedPre)


# Percentage of each ethnicity and universal credit 

PaymentsByRace <- EconData %>%
  filter(!is.na(IndepVar)) %>%
  filter(!is.na(Payments)) %>%
  group_by(IndepVar) %>%
  count(Payments) %>%
  mutate(percPay = n/ sum(n) * 100) %>%
  pivot_wider(c(-n),names_from = IndepVar, values_from = percPay)


PaymentsByRace <- EconData %>%
  filter(!is.na(IndepVar)) %>%
  filter(!is.na(PaymentsBin)) %>%
  group_by(IndepVar, PaymentsBin) %>%
  count(PaymentsBin) %>%
  summarise(n=n(),
    mean=mean(PaymentsBin),
    sd=sd(PaymentsBin)
  ) %>%
  mutate( se=sd/sqrt(n))  %>%
  mutate(percUniCredit = n/ sum(n) * 100) %>%
  pivot_wider(c(-n),names_from = IndepVar, values_from = percUniCredit)


  barchart <- read_excel("Downloads/Boxplot.xlsx")
EconData <- EconData %>% mutate(IndepVar =  as.character(IndepVar))

Boxplot <- ggplot(EconData, aes(x = IndepVar2, y = HourDif)) +
  geom_boxplot(outlier.size=0.5) +
    scale_y_continuous(breaks=seq(-100, 100, by=20)) +
  stat_summary(fun=mean, geom="point", shape=20, size=1, color="red", fill="red") +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", size = 0.1, width = 0.5, colour = "red") +
  labs(x = "Employment type and ethnicity", y = "Difference in working hours") +
  theme_light() 
  
  


##############################################################

##############################################################

# Payments 

 barchart <- read_excel("Downloads/Boxplot.xlsx")
View(Boxplot)
install.packages("Hmisc")
install.packages("tidyverse")
library(tidyverse)

one <- EconData %>%
  group_by(IndepVar2, Payments) %>%
  dplyr::summarise(cases = n()) %>% 
  mutate(pct = cases/sum(cases),
         lbl = scales::percent(pct)) 


df$region <- factor(df$region, levels=c('A', 'E', 'D', 'C', 'B'))



barchart$IndepVar2 <-factor(barchart$IndepVar2, levels = c("Ethnic Minority Employee", "White Employee", "Ethnic Minority Self-Employed", "White Self-Employed"))
  reorder(IndepVar2 = c("Ethnic Minority Employee", "White Employee", "Ethnic Minority Self-Employed", "White Self-Employed"))
  ggplot(barchart, 
       aes(x = IndepVar2,
           y = pct,
           fill = Payments)) + 
  geom_bar(stat = "identity",
           position = "dodge2") +
        geom_errorbar(data = barchart, mapping = aes( ymin = Lower, ymax = Upper), position = position_dodge2(width = 0.5, padding = 0.5), size = 0.1, colour = "red") +
  labs(x = "Employment type and ethnicity", y = "Percentage") +
  theme_light() + 
    scale_y_continuous( breaks = seq(0, 100, 10)) +
    scale_fill_grey()
  

  
# As input for the prop.test() function we need the number of successes x and 
# the number of trials n, i.e. in our case of the number of yes responses and 
# the total number of responses.
## ethnic minority
x = 272
n = 293


# Use this information to create the 95% confidence interval for our sample with 
# the prop.test() function
# Hint: the default confidence level of the prop.test() function is 95%
prop.test(x=x, n=n)

```


Create two new objects with self-employed and employees. 
```{r}
EconData <- EconData %>% mutate(IndepVar = as.character(IndepVar))

Self <- EconData %>%
  filter(IndepVar == 2 | IndepVar == 4) 

Self <- Self %>%
  mutate(IndepVar = ifelse(IndepVar == 2, "EM", "WH"))
         
Self1 <- table(Self$IndepVar, Self$Payments)

Employee <- EconData %>%
  filter(IndepVar == 1 | IndepVar == 3)

EM <- EconData %>%
  filter(IndepVar == 1 | IndepVar == 2)

White <- EconData %>%
  filter(IndepVar == 3 | IndepVar == 4)
```




```{r}
ggplot(EconData, aes(x = factor(IndepVar), y = Payments)) + 
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9)) +
  stat_summary(fun.data = mean_se, geom = "linerange", position = position_dodge(width = .9)) +
  ggtitle("Mean +/- SE")
```

Initial t-test and chi-square tests to show whether the relationships are significant. 

Ethnic minorities and hours worked 
Ha = Ethnic minorities who are self-employed are more likely to face reduced working hours during covid than the equivalent white individual and those who are employees. 
Ho = There is no ethnic occupational difference between those who have and have not had their hours reduced. 
t-test (continuous independent)
```{r}
class(EconData$IndepVar)
## self employed --> does ethnicity make a difference? 
resultSelf1 <- t.test(HourDif ~ IndepVar, Self, alt = "less")
resultSelf

resultSelf1

## employees --> does ethnicity make a difference? 
resultEmpl <- t.test(HourDif ~ IndepVar, Employee, alt = "less")
resultEmpl
## p-value is smaller than the alpha, suggesting that ethnicity causes a relationship between employees and difference in working hours.

```


## EMs and universal credit 
## Ha = Self-employed ethnnic minorities are more likely to apply for universal credit than the equivalent white individual. 
## Ho = There is no occupational ethnic difference between those who have been and hvae not applied for universal credit. 

```{r}

result2Self  <- chisq.test(SelfUni, simulate.p.value = T)

SelfUni

fisher.test(SelfUni)
result2Self$expected

result2Self <- fisher.test(table(Self$IndepVar, Self$UniCredit))
result2Self

result2Empl <- fisher.test(table(Employee$IndepVar, Employee$UniCredit))
result2Empl
## some of the expected counts are less than 5 and one is less than 1, thus result could be invalid and the assumptions have not been met. 
## BUT the p-value is 0.001396, which is less than the alpha of 0.05, suggesting there is a relationship between ethnicity and employment type and universal credit applications. The porability of independent between the two varibales is thus low. 
```

## EMs and late housing payments 
## Ha = Self-employed ethnic minorities are more likely to be behind on housing payments than the equivalent white inidvidual. 
## Ho = There is no ethnic difference betwen those who are and are not behind on housing payments. 

```{r}
result3Self  <- chisq.test(table(Self$IndepVar, Self$Payments))
result3Self
result3Self$expected

result3Empl <- chisq.test(table(Employee$IndepVar, Employee$Payments))
result3Empl

fisher.test(table(Self$IndepVar, Self$Payments))
fisher.test(table(Employee$IndepVar, Employee$Payments))

fisher.test(Self1)
```


```{r}
install.packages("writexl")
library(writexl)

write_xlsx(one, "~/Downloads/Boxplot.xlsx")

x <- c('tidyverse', 'openxlsx')
sapply(X=x, FUN = require, character.only = TRUE)

# Create graph---------
dcs <- ggplot(mtcars) +
  geom_point(aes(cyl, disp))

# Export Pic -----------
pic_path <- "~/Downloads/dcs.jpeg"
png(filename = pic_path)
plot(Boxplot)
dev.off()

library(openxlsx)
# Add to a new work book -------------
wb <- openxlsx::createWorkbook()
addWorksheet(wb, "Plots")
insertImage(wb, "Plots", pic_path)
openxlsx::saveWorkbook(wb,
                       "~/example.xlsx")

# Kill pic
unlink(pic_path)
```

